### 2018-02-01 includes/appendix.tex
appendix

In this section, we will discuss how we proved equisatisfiability guarantees
that transformations provide in \fuzzer{}.

We assume that the reader is familiar with SMT-LIB syntax. The transformations
work over boolean, integer, string literals and expressions, and regular
expressions.

To shorten our proofs, we will make the following simplifications:
\begin{enumerate}
  \item
    Since many of the operators can be expressed in terms of other, our
    proofs will only consider a subset of the operators. In particular we will
    drop the following:
    \begin{itemize}
      \item \texttt{||} (boolean or)
      \item \texttt{=} for boolean expressions
      \item \texttt{<=}, \texttt{>}, \texttt{>=}
      \item \texttt{str.prefixof}, \texttt{str.suffixof}, \texttt{str.at}, \texttt{str.contains}
      \item \texttt{re.range}, \texttt{re.+}, \texttt{re.allchar}, \texttt{re.all}
    \end{itemize}

  \item
    In the above, we assume that \texttt{str.at} is expressible using
    \texttt{substring} although the semantics slightly differ when the index is
    out of bounds.

  \item
    We will assume that our program instances are a single assertion. Multiple
    assertions can be expressed by combining them using \texttt{and}.
\end{enumerate}

\begin{definition}
  A \emph{model} is a mapping from boolean, integer, and string variables to
literals of the corresponding type.
\end{definition}

\begin{definition}
  A \emph{program} $P$ satisfies a model $m$ if $P$ evaluates to \texttt{true}
under the model $m$. The function $\eval (P)$ evaluates a program instance
under a model; the specific model used is left implicit as it is clear from
context.
\end{definition}

The proofs all follow a very similar structure. For each transformation
\textit{Trans}, we show that a problem instance $P$ is satisfiable by a model $m$
if and only if $\textit{Trans}(P)$ is satisfiable by $\textit{Trans}(m)$, where
$\textit{Trans}(m)$ is the transformation applied to all the constants in the
model $m$.

To show the above, we show that for all expressions $e$, $\textit{Trans}(\eval
(e)) = \eval (\textit{Trans}(e))$. This is proven using a straightforward
induction on the structure of expressions. Since the transformations fix
boolean constants, we have, $\textit{Trans}(\eval (P)) = \eval (P)$, and hence
if the original problem is satisfiable then the transformed problem is also
satisfiable.

For transformations that have an inverse, such as \textit{Translate} and
\textit{Reverse} this also shows that if the transformed problem is satisfiable,
then the original problem is satisfiable, proving that they are equisat
transformations. \textit{Multiply} does not have an inverse and hence it only
guarantees to take satisfiable problems to satisfiable problems. This can be
seen in the problem instance $y<x<y+1$ which is UNSAT, but multiplying by 2
transforms the problem into $y<x<y+2$ which is SAT.

\subsection{Translate}
We now prove that \textit{Translate} transforms programs into equisatisfiable
programs. From above, we only need to show that for every expression $e$,
$\textit{Translate}(\eval (e)) = \eval (\textit{Translate}(e))$. For now, we
assume that \textit{Translate} fixes the digit characters pointwise and we prove
the stronger fact that $\textit{Translate}$ also fixes the digits pointwise in
any string expression.

@@

\subsection{Reverse}
\textit{Reverse} does not transform programs that use \texttt{str.replace},
\texttt{str.indexof}, \texttt{str.to.int}, and \texttt{str.from.int}. These
operators do not have a well-defined notion of a reverse. The required lemma is
proved by induction on the rest of the operators similar to proof for
\textit{Translate}.

We end our discussion on \textit{Reverse} by nothing that although it disallowed
\texttt{str.replace}, it only did so because it didn't have a natural notion of
a reverse in the theory we were working with. If we expand our theory to also
include an operator that replaces the last instance, we can transform problems
that use \texttt{str.replace}.

\subsection{Multiply}
\textit{Multiply} does not transform programs that use \texttt{*},
\texttt{str.indexof}, \texttt{str.to.int}, and \texttt{str.from.int}. These
operators do not have a well-defined notion of a multiply. The required lemma is
proved by induction on the rest of the operators similar to the above proofs.

>>>>

In this section, we discuss how we proved equisatisfiability
guarantees that transformations provide in \fuzzer{}. We assume that
the reader is familiar with SMT-LIB syntax. The transformations work
over boolean, integer, string literals and expressions, and regular
expressions.

To shorten our proofs, we make the following simplifications:
\begin{enumerate}
  \item
    Since many of the operators can be expressed in terms of others,
    our proofs will only consider a subset of the operators. In
    particular we drop the following:
    \begin{itemize}
      \item \texttt{||} (boolean or)
      \item \texttt{=} for boolean expressions
      \item \texttt{<=}, \texttt{>}, \texttt{>=}
      \item \texttt{str.prefixof}, \texttt{str.suffixof}, \texttt{str.at}, \texttt{str.contains}
      \item \texttt{re.range}, \texttt{re.+}, \texttt{re.allchar}, \texttt{re.all}
    \end{itemize}

  \item
    In the above, we assume that \texttt{str.at} is expressible using
    \texttt{substring} although the semantics slightly differ when the index is
    out of bounds.
\end{enumerate}

\begin{definition}
  A \emph{model} is a mapping from boolean, integer, and string variables to
elements in the domain of universe of the corresponding type.
\end{definition}

\begin{definition}
  A \emph{formula} $P$ satisfies a model $m$ if $P$ evaluates to
  \texttt{true} under the model $m$. The function $\eval (P)$
  evaluates a formula under a model; the specific model used is left
  implicit as it is clear from context.
\end{definition}

The proofs all follow a very similar structure. For each
transformation \textit{Trans}, we show that a formula $P$ is
satisfiable by a model $m$ if and only if $\textit{Trans}(P)$ is
satisfiable by $\textit{Trans}(m)$, where $\textit{Trans}(m)$ is the
transformation applied to all the constants in the model $m$. (The
exception here is the \textit{Multiply} operator as discussed below.)

To show the above, we show that for all expressions $e$,
$\textit{Trans}(\eval (e)) = \eval (\textit{Trans}(e))$. This is
proven using a straightforward induction on the structure of
expressions. Since the transformations fix boolean constants, we have,
$\textit{Trans}(\eval (P)) = \eval (P)$, and hence if the original
problem is satisfiable so is the transformed problem.

For transformations that have an inverse, such as \textit{Translate}
and \textit{Reverse} this also shows that if the transformed problem
is satisfiable, then the original problem is satisfiable, proving that
they are equisat transformations. \textit{Multiply} (as defined here)
does not have an inverse and hence it only guarantees to take
satisfiable problems to satisfiable problems. This can be seen in the
problem instance $y<x<y+1$ which is UNSAT, but multiplying by 2
transforms the problem into $y<x<y+2$ which is SAT.

\subsection{Translate}
We now prove that \textit{Translate} transforms constraints into
equisatisfiable constraints. From above, we only need to show that for
every expression $e$, $\textit{Translate}(\eval (e)) = \eval
(\textit{Translate}(e))$. For now, we assume that \textit{Translate}
fixes the digits pointwise and we prove the stronger fact that
$\textit{Translate}$ also fixes characters pointwise in any string
expression. (I.e., $\textit{Translate}$ are bijections and further
that digits are not mapped to any other character.)

@@

\subsection{Reverse}
\textit{Reverse} does not transform constraints that use
\texttt{str.replace}, \texttt{str.indexof}, \texttt{str.to.int}, and
\texttt{str.from.int}. These operators do not have a well-defined
notion of a reverse. The required lemma is proved by induction on the
rest of the operators similar to proof for \textit{Translate}.

We end our discussion on \textit{Reverse} by noting that although it
disallowed \texttt{str.replace}, it only did so because it didn't have
a natural notion of a reverse in the theory we were working with.
%% If we expand our theory to also include an operator that replaces
%% the last instance, we can transform problems that use
%% \texttt{str.replace}.

\subsection{Multiply}
\textit{Multiply} does not transform constraints that use \texttt{*},
\texttt{str.indexof}, \texttt{str.to.int}, and
\texttt{str.from.int}. These operators do not have a well-defined
notion of a multiply. The required lemma is proved by induction on the
rest of the operators similar to the above proofs.
